---
title: "Cesar R workshop 2: welcome to the tidyverse "
author: "Alex Slavenko & Evatt Chrigwin"
output: html_notebook
date: "`r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
fontsize: 12pt
output:
  pdf_document:
    toc: true
    toc_depth: 2
urlcolor: blue
linkcolor: red

---
## Task 1.
### Load allfollowing packages: `tidyr`, `dplyr`, `magrittr`, `readr`, `ggplot2`?

#You should have all of these packages installed from the previous workshop, but if not install them using install.package("package_name"). 

```{r}
#load packages
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)

```


## Task 2.
### Set your working directory for today
```{r}

setwd('/Users/Evatt/Dropbox (cesar & EnviroDNA)/Evatt/R_workshop')

```

## Task 3.
### Import the  dataset (workshop_2_data.csv) and store it as an object named data
```{r}
data1<-read.csv('Play_data.csv')
```

## Task 4.
### Have a look at your data
### Is your data long or wide?
```{r}
head(data1)
str(data1)
```


## Task 5.
### Surivial was scored at multiple times during the experiment.Today, we will only consider survival at the last scoring time
### Idetify the latest scoring time using distinct() and then use filter() to create a new dataset that exludes all other scoring times
```{r}
distinct(data1,Time)
data2<-filter(data1, Time=='72h')
str(data2)

table(data1$Population)
```
## Task 6.
### Raw data often more infomation that what we need (e.g., extra traits and general notes) that can make things messy
### Today, we are only interested in 5 coloumns: Population, Rep, Time, Alive, Dead 
### Can you use the select() to create a new dataset only includes these 5 coloumns? 
```{r}
data3<-select(data2,'Population', 'Rep', 'Time', 'Alive', 'Dead')
head(data3)
```

## Task 7.
### We are interested in the survival rate of our critters but we only have the number alive and dead per replicate
### We can use the mutate() to create new dataset with the coloumn for survival rate

```{r}
data4<-mutate(data3, Survivial=Alive/(Alive+Dead))
```

## Task 8.
### Tasks 4-6 have allowed us create our "Final" dataset that is ready for analysis. However, doing so also required us to make 3 new and seperate dataset along the way... All these extra datasets can add up and create confusion. Luckily there is a better way!

### We can use pipes "%>%" to complete tasks 4-6 in a single line of code.  
######## NOTE EVATT NEEDS TO MAKE SLIDES TO EXPLAIN PIPES #####
```{r}
data5<-data1%>%
  filter(Time=='72h')%>%
  select('Population', 'Rep', 'Time', 'Alive', 'Dead')%>%
  mutate(Survivial=Alive/(Alive+Dead))
```

## Task 9.
### Let's get some answers from our data!
### Find the  average survival using summerise()
```{r}
data5%>%
  summarise(mean(Survivial))
```



## Task 10.
### Now we want to know the average survival for each population
### Calculate the average suvivial of each population by using  piping to add with group_by() to code above.
```{r}
data5%>%
  group_by(Population)%>%
  summarise(mean(Survivial))
```


###Task 10
### using count(), Find the total number of replicates of each population and the total number of individuals across all replicates

```{r}
data5%>%
    group_by(Population)%>%
    count()
```

### using summerise(), Find the total number of individuals tested within each population

```{r}
data5%>%
    group_by(Population)%>%
    summarise(total=sum(Alive+Dead))
```


## Task 11.
### Now we have the data all neat and tidy, let's save this version and send to send to other people.
```{r}
write.csv(data5, 'cool_old_data.csv')

```



## Task 12
### Oh cool someone sent us some new data! Let's add this to our existing dataset!
### Import the data
```{r}
data_new<-read.csv('Oh_cool_some_new_data.csv')
```

## Task 4.
### Have a look at your data
### Is your data long or wide?
```{r}
head(data_new)
str(data_new)
```
## Task 13.
### the new data has a different format (wide) than your old data (narrow)
### before we can combine them, we need them to be in the same format
```{r}
data_size<-data_new %>% pivot_longer(cols=c('Brunswick',	'Carlton','Coburg',	'Northcote'),
                          names_to = "Population" , 
                          values_to = "Growth")

```


## Task 14.
### combine datsets using join
```{r}
df_fin <- left_join(data5, data_size, by=c('Population', 'Rep'))
```

## Task 15.
### plot(if we have time)
```{r}
df_fin%>%
ggplot(aes(x=Growth, y=Survivial)) + geom_point()
```

